Brief Announcement: Tight Bounds for Online Balanced Repartitioning

### 2020-05-21 podc_NEW/podc_ba.tex
wip

\begin{abstract}
	Distributed   applications,  including  batch  processing, streaming, scale-out databases,
	or machine learning, generate a~significant amount of network traffic.
	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
	However, the communication pattern of different applications is often unknown a~priori and may change over time, hence it needs to be learned in an~online manner.
	%
	This paper revisits the online 
	balanced partitioning problem 
	(introduced by Avin et al.~at DISC 2016)
	that asks for an algorithm that strikes
	an optimal tradeoff between the benefits
	of collocation (i.e., lower network load) 
	and its costs (i.e., migrations). 
	%
	Our first contribution is a~significantly improved deterministic
	lower bound of $\Omega(k\cdot \ell)$ on the
	competitive ratio, where $\ell$ is the number
	of clusters and $k$ is the cluster size,
	even for a~scenario in which the communication
	pattern is static and can be perfectly partitioned;
	we also provide an asymptotically tight upper bound 
	of $O(k\cdot \ell)$ for this scenario.
	For $k=3$, we contribute an asymptotically tight upper bound
	of $\Theta(\ell)$
	for the general model in which the
	communication pattern can change arbitrarily over time.
	We improve the result for $k=2$ by providing a~strictly $6$-competitive upper bound for the general model.
	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.

@@

\noindent \textbf{Contributions.}
%
This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
For the learning model, we present a~lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
(that holds also in the general partitioning model).
The best known lower bound so far was $\Omega(k)$~\cite{repartition-disc} that holds only in the general partitioning model.
We complement this result with an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the learning model.
For the general partitioning model, we design
an asymptotically optimal,
$\Theta(\ell)$-competitive algorithm for $k=3$, improving the best known upper bound 
so far $O(\ell^2)$~\cite{repartition-disc}.
We further present a~strictly $6$-competitive algorithm for $k=2$ that improves upon the previous $7$-competitive algorithm with $O(\alpha\ell^2)$ additive constant.

@@

We provide a~lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
The lower bound requires $k\geq 3$.
In contrast, for $k=2$ the learning problem is trivial: immediate collocation of communicating pairs is $1$-competitive.
%, a~lower bound of $3$ is known~\cite{repartition-disc}, and we provide a~$6$-competitive algorithm, see Section~\ref{sec:k2}.)

  \begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
	 $(k-2)(\ell-1)/2 - 2$ for any $k\geq 3$ and $\ell \geq 2$.
\end{theorem}

A~\emph{ground set} is a group of nodes that communicate if they are not collocated.
The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
We start by constructing a~ground set $B$ of size $k-1$ on an arbitrarily chosen cluster.
In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
%, and we refer to it as the \emph{pivot} node.
We issue requests between this node and some node that was initially collocated with it.
By repeating such requests, almost every node is once collocated with the ground set $B$.
The cost of the algorithm is linear in terms of the number of nodes, and there exists an optimal offline algorithm \OPT that performs only two node exchanges.

@@

There exists an asymptotically optimal algorithm for the learning problem.
The algorithm  collocates  a~pair as soon as they communicate and it never separates them.
To preserve collocated pairs,
we employ the concept of components,
introduced by Avin~et~al.~\cite{repartition-disc}.

@@

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of serving each request is at most $2\cdot\OPT$.
\end{lemma}

@@

\begin{theorem}	\label{thm:upperbound}
	\PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{theorem}

\section{General Partitioning Model}
\label{sec:part}

Now  we discuss the general online
model where the request sequence
can be arbitrary.
In Section~\ref{sec:k3}, we show an $O(k \cdot \ell)$-competitive algorithm for $k=3$ using the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental}.

\subsection{Optimal Algorithm for Clusters of Size 3}
\label{sec:k3}

The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, which for $k=3$ is $O(\ell^2)$-competitive.
In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.
This allows to bound the cost of repartition by a constant (Lemma~\ref{lem:1req}).

@@

\noindent
\textbf{Component-based algorithm.}
The algorithm \TAlg partitions nodes into components, and
initially, each node is isolated (belongs to its own component).
For each pair of nodes $\set{x,y}$, \TAlg maintains a~counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to singleton components, resets all counters to $0$, and ends the phase.

\begin{lemma}
	\label{lem:1req}
	In a~single repartition of nodes (after a~merge of components), \TAlg exchanges at most two pairs of nodes.
\end{lemma}

\begin{theorem}
	\TAlg is $60\ell$-competitive for $k=3$.
	\label{thm:k=3}
\end{theorem}

>>>>

\begin{abstract}
	Distributed   applications,  including  batch  processing, streaming, scale-out databases,
	or machine learning, generate a~significant amount of network traffic.
	By collocating frequently communicating nodes (e.g., virtual machines) on the same clusters (e.g., server or rack), we can reduce the network load and  improve application performance. 
	However, the communication pattern of different applications is often unknown a~priori and may change over time, hence it needs to be learned in an~online manner.
	%
	This paper revisits the online 
	balanced partitioning problem 
	(introduced by Avin et al.~at DISC 2016)
	that asks for an algorithm that strikes
	an optimal tradeoff between the benefits
	of collocation (i.e., lower network load) 
	and its costs (i.e., migrations). 
	%
	Our first contribution is a~significantly improved deterministic
	lower bound of $\Omega(k\cdot \ell)$ on the
	competitive ratio, where $\ell$ is the number
	of clusters and $k$ is the cluster size,
	even for a~scenario in which the communication
	pattern is static and can be perfectly partitioned;
	we also provide an asymptotically tight upper bound 
	of $O(k\cdot \ell)$ for this scenario.
	For $k=3$, we contribute an asymptotically tight upper bound
	of $\Theta(\ell)$
	for the general model in which the
	communication pattern can change arbitrarily over time.
	In contrast to most prior work, our algorithms respect all capacity constraints and do not require resource augmentation.

@@

\noindent \textbf{Contributions.}
%
This paper presents several new results on the dynamic graph partitioning problem  without augmentation.
For the learning model, we present a~lower bound of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
(that holds also in the general partitioning model).
The best known lower bound so far was $\Omega(k)$~\cite{repartition-disc} that holds only in the general partitioning model.
We complement this result with an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the learning model.
For the general partitioning model, we design
an asymptotically optimal,
$\Theta(\ell)$-competitive algorithm for $k=3$, improving the best known upper bound 
so far $O(\ell^2)$~\cite{repartition-disc}.

@@

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for the learning model of Dynamic Balanced Graph Partitioning is at least
	 $\Omega(k\cdot \ell)$ for any $k\geq 3$ and $\ell \geq 2$.
\end{theorem}

We choose a set $B$ of $k-1$ collocated nodes and issue requests between them so they must stay collocated.
In any partition, there must exist an isolated node collocated with $B$.
%, and we refer to it as the \emph{pivot} node.
We issue requests between this node and some node that was initially collocated with it.
We repeat such requests to the current node collocated with $B$, and this node get replaced by another node that was never requested.
We issue $(k-2)(\ell - 1)-4$ requests this way.
The cost of the algorithm is linear in terms of the number of nodes, and there exists an optimal offline algorithm \OPT that performs only two node exchanges.

@@

We present the algorithm \PPL that is  asymptotically optimal for the learning problem.
It utilizes the concept of components,
introduced by Avin~et~al.~\cite{repartition-disc}.

@@

@@

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of a single repartition is at most $2\cdot\OPT$.
	The corollary is that \PPL is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{lemma}

\section{General Partitioning Model: Optimal Algorithm for Clusters of Size 3}

\label{sec:part}
\label{sec:k3}

Now  we discuss the general online
model where the request sequence
can be arbitrary.
The algorithm analyzed in this section is a modified version of the algorithm DET proposed by Avin et al.~\cite{repartition-disc}, which for $k=3$ is $O(\ell^2)$-competitive.
In our algorithm, we choose the closest partition after a component merge instead of an arbitrary one.
This allows to bound the cost of repartition by a constant:
\begin{lemma}
	\label{lem:1req}
	In a~single repartition of nodes, the algorithm exchanges at most two pairs of nodes.
\end{lemma}

@@

\begin{theorem}
	The proposed algorithm for $k=3$ is $O(\ell)$-competitive.
	\label{thm:k=3}
\end{theorem}
